Precision calculations of the cosmic shear power spectrum projection

### 2017-02-17 limber2.tex
Added in some instances of "equation"~\ref{} for MNRAS style

We compute the spherical-sky weak-lensing power spectrum of the shear and
convergence. We discuss various approximations, such as flat-sky, and first- and
second-order Limber equations for the projection. We find that the impact of
adopting these approximations are negligible when constraining cosmological
parameters from current weak lensing surveys.
%
This is demonstrated using data from the Canada-France-Hawaii Lensing Survey
(CFHTLenS). We find that the reported tension with Planck Cosmic Microwave
Background (CMB) temperature anisotropy results cannot be alleviated, in
contrast to the recent claim made by Kitching et al. (2016, version 1)

@@

On a sphere, correlation functions formally cannot be related to the power
spectrum by the Hankel transform in equation~(\ref{eqn:xiGG}), and should
be replaced by the \mk{spherical} transform \citep{1973ApJ...185..413P}
%
\begin{equation}
 \xi_+(\theta) = \mk{\left\langle \gamma \gamma^\ast \right\rangle(\theta)}
  =  \frac 1{4\pi} \sum\limits_{\ell=2}^\infty (2 \ell + 1) C^\gamma(\ell) {\rm P}_\ell[\cos(\theta)],
  \label{eq:xi_Legendre}
\end{equation}
%
where ${\rm P}_\ell(x)$ are the Legendre polynomials in $x$ of order $\ell$. \mk{This equation is easily derived
using the harmonics expansion of the shear (\ref{eq:gamma_harm_exp}) and the orthogonality relation of
the spherical harmonics (\ref{eq:Yellm_ortho}).}

The spherical power spectrum is formally not defined for non-integer $\ell$
\citep[see][for alternative spherical-sky formulae for the two-point
correlation function]{2005PhRvD..72b3516C}, as functions defined on the
sphere are necessarily periodic.
As we have shown in section~\ref{sec:comp}, however, the spherical second-order
extended Limber approximation provides a percent-level precision representation
of the full spherical projection for $\ell > 3$.
The spherical pre-factor $\ellbar(\ell, 2)$ (equation~\ref{eq:ellbar}) can be
generalised to non-integer arguments and is positive for $\ell \ge 2$. We can
thus use the spherical power spectrum \sout{in the Limber approximations} with
the Hankel transformation in equation~(\ref{eqn:xiGG}) to compute the two-point correlation
functions. This has the advantage being able to use very fast numerical implementations of
the Hankel tranforms using FFTs \citep{fftlog}, when Monte-Carlo sampling.

We compare predictions for the two-point shear correlation function using
the Hankel transformation and the full spherical transformation in
Fig.~\ref{fig:xi_p_legendre}. We show the full projection and the best
approximation, ExtL2Sph. Note that for the `FullSph' case we replace the full
projection with ExtL2Sph for $\ell > 200$ to reduce computation time.
%
\ch{We find that the Hankel transform (equation~\ref{eqn:xiGG}) is accurate to better than $2$ percent level.
%
The difference between second-order Limber and full projection using the
spherical transform (\ref{eq:xi_Legendre}) is well below one percent
for $\vartheta < 200$ arcminutes.}
The red circles in Fig.~\ref{fig:xi_p_legendre} present the limit of
precision that we can achieve with the current fast Hankel transform implementation
for the correlation function. The green line shows the limit of the
second-order Limber approximation on the correlation function.

>>>>

We compute the spherical-sky weak-lensing power spectrum of the shear and
convergence. We discuss various approximations, such as flat-sky, and first- and
second-order Limber equations for the projection. We find that the impact of
adopting these approximations are negligible when constraining cosmological
parameters from current weak lensing surveys.
%
This is demonstrated using data from the Canada-France-Hawaii Telescope Lensing Survey
(CFHTLenS). We find that the reported tension with Planck Cosmic Microwave
Background (CMB) temperature anisotropy results cannot be alleviated, in
contrast to the recent claim made by Kitching et al. (2016, version 1)

@@

On a sphere, correlation functions formally cannot be related to the power
spectrum by the Hankel transform in equation~(\ref{eqn:xiGG}), and should
be replaced by the \mk{spherical} transform \citep{1973ApJ...185..413P}
%
\begin{equation}
 \xi_+(\theta) = \mk{\left\langle \gamma \gamma^\ast \right\rangle(\theta)}
  =  \frac 1{4\pi} \sum\limits_{\ell=2}^\infty (2 \ell + 1) C^\gamma(\ell) {\rm P}_\ell[\cos(\theta)],
  \label{eq:xi_Legendre}
\end{equation}
%
where ${\rm P}_\ell(x)$ are the Legendre polynomials in $x$ of order $\ell$. \mk{This equation is easily derived
using the harmonics expansion of the shear (equation~\ref{eq:gamma_harm_exp}) and the orthogonality relation of
the spherical harmonics (equation~\ref{eq:Yellm_ortho}).}

The spherical power spectrum is formally not defined for non-integer $\ell$
\citep[see][for alternative spherical-sky formulae for the two-point
correlation function]{2005PhRvD..72b3516C}, as functions defined on the
sphere are necessarily periodic.
As we have shown in section~\ref{sec:comp}, however, the spherical second-order
extended Limber approximation provides a percent-level precision representation
of the full spherical projection for $\ell > 3$.
The spherical pre-factor $\ellbar(\ell, 2)$ (equation~\ref{eq:ellbar}) can be
generalised to non-integer arguments and is positive for $\ell \ge 2$. We can
thus use the spherical power spectrum with
the Hankel transformation in equation~(\ref{eqn:xiGG}) to compute the two-point correlation
functions. This has the advantage being able to employ fast FFT numerical implementations of
the Hankel tranforms \citep{fftlog}, when Monte-Carlo sampling.

We compare predictions for the two-point shear correlation function using
the Hankel transformation and the full spherical transformation in
Fig.~\ref{fig:xi_p_legendre}. We show the full projection and the best
approximation, ExtL2Sph. Note that for the `FullSph' case we replace the full
projection with ExtL2Sph for $\ell > 200$ to reduce computation time.
%
\ch{We find that the Hankel transform (equation~\ref{eqn:xiGG}) is accurate to better than $2$ percent level.
%
The difference between the second-order Limber and full projection solution using the
spherical transform (equation~\ref{eq:xi_Legendre}) is well below one percent
for $\vartheta < 200$ arcminutes.}
The red circles in Fig.~\ref{fig:xi_p_legendre} present the limit of
precision that we can achieve with the current fast Hankel transform implementation
for the correlation function. The green line shows the limit of the
second-order Limber approximation on the correlation function.
